---
title: "SeeMe Scotland"
author: "Shahrukh Gill"
date: "04/10/2020"
output: html_document
---

'Base R' is the Original language used in R and is always where we start.

when begining it is important to create working directory.
click the "Session" tab, then "set working directory" and then "Choose Directory..." to find a suitable place to keep all the files we make.


Base R can be used to create data frames, vectors and matrixes.

Here are some vectors we can create
```{r}
c(2, 4, 6) 

2:6 

seq(2, 3, by=0.5)

rep(1:2, times=3)

rep(1:2, each=3)
```



Here is a random data frame we can make.
```{r}
mydata <- data.frame(id = 1:5, x = c('stigma', 'mental health', 'discrimination', 'wellbeing', 'engagment'), Time = c(5.50,3.23,11.23,1.15,10.59), n = rep(50,5), y = c(20,35.4,45,55.5,70.6), z = c(6,17,26,37,44))

mydata
```

Hashtags"#" can help when you want to leave comments or harmless data inside the code.

```{r}
#try and debug this to match the original code and create the "mydata" data frame
4*5
```

Base R can also write data
```{r}
write.csv(mydata, "mydata.csv")
```
and read data
```{r}
mydata2 <- read.csv("mydata.csv") 
```

R also has pachages like tidyverse and ggplot2 and several others that save your time and energy.

Lets have a look at one. 

They may need to be insalled if you have not installed them before using this function install.packages() e.g. install.packages("tidyverse")

```{r}
library(tidyverse)
library(ggplot2)
library(psych)
```

```{r}

```

also, Base R usues conditions which can be used inside functions:
a == b Are equal 
a > b Greater than 
a >= b Greater than or equal to 
is.na(a) Is missing
a != b Not equal 
a < b Less than 
a <= b Less than or equal to 
is.null(a) Is null
